\documentclass[10.3pt, a4paper ]{article}
\frenchspacing

\input{util/packages}
\input{util/commands}

\input{sections/title}

\begin{document}

% \sloppy

\maketitle
\input{sections/abstract.tex}
% \newpage
% \tableofcontents
% \newpage
\begin{multicols}{2}

\input{sections/introduction}
\input{sections/results.tex}
\input{sections/discussion.tex}
\input{sections/methods.tex}


\subsection*{Data Availability}\label{sec:data_availability}
All the data and simulations presented in this work are available from the corresponding author upon request. Code used in this work is available from the corresponding author upon request, or on Github \cite{briancoyle_implementation_2019}.
\subsection*{Acknowledgements}\label{sec:acknowledgements}

B.C.\@ thanks Andru Gheorghiu for useful discussions and title suggestion. We also thank Jean Feydy, Patric Fulop, Vojtech Havlicek and Jiasen Yang for clarifying pointers. We also thank Atul Mantri for comments on the manuscript. This work was supported by the Engineering and Physical Sciences Research Council (grant EP/L01503X/1), EPSRC Centre for Doctoral Training in Pervasive Parallelism at the University of Edinburgh, School of Informatics and Entrapping Machines, (grant FA9550-17-1-0055). We also thank Rigetti Computing for the use of their Quantum compute resources, and views expressed in this paper are those of the authors and do not necessarily reflect the views or policies of Rigetti Computing.

\subsection*{Author Contributions}
B.C.\@ devised the theoretical aspects of the work, and wrote the code for the numerical results with the help of D.M.. D.M.\@ contributed to the Learning Supremacy definition. 
E.K.\@ and V.D.\@ supervised the work. All authors contributed to the manuscript writing. 

\subsection*{Competing Interests}
The authors declare no conflict of interests.

% \addbibresource{isingborn.bib}
% \printbibliography
\bibliography{npj_main}
\bibliographystyle{naturemag}
\end{multicols}


\appendix
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thefigure}{S\arabic{figure}}

\setcounter{figure}{0}
\newpage
\section*{Supplementary Material}
\input{sections/supp_matt_ibm_training.tex}
\input{sections/supp_matt_kernel_methods.tex}
\input{sections/supp_matt_stein_discrepancy.tex}
\input{sections/supp_mat_sinkhorn_sample.tex}
\input{sections/supp_matt_extra_numerics.tex}
\input{sections/supp_matt_supremacy.tex}
\input{sections/supp_mat_learning_supp.tex}

\bibliographyS{supp_matt}
\bibliographystyleS{naturemag}


\end{document}
